---
title: Hospitalization Parameterized report
output: 
  html_document:
    code_folding: hide
params:
  forecasters:
    label: "Forecasters:"
    choices: [Covid19Sim-Simulator, CovidActNow-SEIR_CAN, COVIDhub-ensemble, COVIDhub-trained_ensemble, CU-select, Google_Harvard-CPF, GT-DeepCOVID, IHME-CurveFit, JHU_IDD-CovidSP, JHUAPL-Bucky, JHUAPL-Gecko, JHUAPL-SLPHospEns, Karlen-pypm, LANL-GrowthRate, MOBS-GLEAM_COVID, UCLA-SuEIR, UCSB-ACTS, USACE-ERDC_SEIR, USC-SI_kJalpha]
    value: [Covid19Sim-Simulator, COVIDhub-ensemble, COVIDhub-trained_ensemble, CU-select, JHU_IDD-CovidSP, Karlen-pypm]
    input: select
    multiple: TRUE
  primary_forecaster:
    label: "Primary forecaster:"
    choices: [Covid19Sim-Simulator, CovidActNow-SEIR_CAN, COVIDhub-ensemble, COVIDhub-trained_ensemble, CU-select, Google_Harvard-CPF, GT-DeepCOVID, IHME-CurveFit, JHU_IDD-CovidSP, JHUAPL-Bucky, JHUAPL-Gecko, JHUAPL-SLPHospEns, Karlen-pypm, LANL-GrowthRate, MOBS-GLEAM_COVID, UCLA-SuEIR, UCSB-ACTS, USACE-ERDC_SEIR, USC-SI_kJalpha]
    value: COVIDhub-ensemble
    input: select  
  colorblind_palette: TRUE
  weeks: 
    label: "Weeks"
    value: [1,2,3,4]
    input: select
    choices: [1,2,3,4]
    multiple: True
  printcode: TRUE
editor_options: 
  chunk_output_type: console
---

**Package installation**
```{r common-stuff, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(autodep = TRUE, fig.align = "center", message = FALSE, warning = FALSE, echo=params$printcode)
library(covidcast)
library(evalcast)
library(modeltools)
#library(quantgen)
library(tidyverse)
library(zookeeper)
library(namespace)
#library(devtools)
#library(usethis)
library(downloadthis)
library(assertthat)  # needed for this particular forecaster
library(sf)
library(Cairo)
library(tidyverse)
library(viridis)
library(ggplot2)
library(plotly)
library(here)
```

```{r, results = FALSE}
aheads  <- as.numeric(params$weeks)
url_hosp <- "https://forecast-eval.s3.us-east-2.amazonaws.com/score_cards_state_hospitalizations.rds"
download.file(url_hosp, "eval_hospitalizations.RDS") # download to disk
scores1 <- readRDS(paste0(here(), "/eval_hospitalizations.RDS"))
scores <- subset(scores1,  forecaster %in% params$forecasters)
```

```{r get-forecast-dates, results = FALSE}
our_pred_dates <- 
  scores %>%
  filter(forecaster == "COVIDhub-ensemble") 

our_pred_dates <- unique(our_pred_dates$forecast_date) 
n_dates <- length(our_pred_dates)

# n_dates - 4 is often the most recent date with ground truth for 4 weeks ahead
# dates spaced out 2 weeks ahead to see different behaviors
forecast_dates <- our_pred_dates[n_dates- 2 *5:2]
```


```{r,results = FALSE}
scores %>%
  filter(forecaster =="Karlen-pypm")

scores$forecast_date <- 
  if_else(scores$forecaster %in% c("Karlen-pypm", "CU-select","Covid19Sim-Simulator","JHU_IDD-CovidSP"), scores$forecast_date + 1, scores$forecast_date)

scores <- subset(scores, forecast_date %in% forecast_dates & ahead <= aheads)
results <- 
  scores %>% 
  select(c("ahead", "geo_value", "forecaster","forecast_date", "data_source", "signal","target_end_date","incidence_period","actual","wis","ae","cov_80"))

results %>%
  group_by(forecast_date) %>%
  summarise(n_distinct(geo_value))
```

## Downloading the data

Retrieving prediction data of the selected forecasters

```{r download data}
results %>%
  download_this(
    output_name = "results",
    output_extension = ".csv",
    button_label = "Download Hospital Evaluation Dataset",
    button_type = "success",
    has_icon = TRUE,
    csv2 = FALSE,
    icon = "fa fa-save"
  )
```

## Overall AE, WIS, Coverage 80 {.tabset}
NOTE: Results are based on the following numbers of common locations

### Overall Absolute Error
By Weeks Ahead
```{r overall, height=800, width= 1000}
weeks.label = c("1 week ahead", "2 weeks ahead", "3 weeks ahead", "4 weeks ahead")
names(weeks.label) = c(1, 2, 3, 4)

subtitle = sprintf("Forecasts made over %s to %s",
                   format(min(forecast_dates), "%B %d, %Y"),
                   format(max(forecast_dates), "%B %d, %Y"))

plot_ae <-
  plot_canonical(results, 
                 x = "ahead", 
                 y = "ae", 
                 aggr = mean) +
  labs(title = subtitle, x= "Weeks Ahead", y = "Mean AE",color='Forecasters') +  
  geom_point(aes(text=sprintf("Weeks Ahead: %s<br>Average Error: %s <br>Forecaster: %s", 
                              ahead, 
                              round(ae, digits=2),
                              color)),
             alpha = 0.05) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_log10()

if (params$colorblind_palette) {
  plot_ae <- plot_ae + 
    scale_color_viridis_d()
}

ggplotly(plot_ae, tooltip="text", width=1000) %>% 
  layout(hoverlabel = list(bgcolor = "white"))
```

### Overall Weighted Interval Score
By Forecast Dates
``` {r}
plot_wis <-
  plot_canonical(results, 
                 x = "forecast_date", 
                 y = "wis", 
                 aggr = mean,
                 grp_vars = c("forecaster","ahead"), 
                 facet_rows = "ahead") + 
  labs(title = subtitle, 
       x = "Forecast Dates", 
       y = "Mean WIS",
       color = "Forecasters") +
  geom_point(aes(text=sprintf("Forecast Date: %s<br>Mean WIS: %s <br>Forecaster: %s", 
                              forecast_date, 
                              round(wis, digits = 2),
                              color)),
             alpha = 0.05) +
  facet_wrap(~ahead, nrow = 4, labeller = labeller(ahead=weeks.label)) +
  theme(strip.background = element_rect(fill = "white")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = "center")) + 
  scale_y_log10()

if (params$colorblind_palette) {
  plot_wis <- plot_wis + 
    scale_color_viridis_d()
}

ggplotly(plot_wis, tooltip="text", height=800, width= 1000) %>% 
  layout(hoverlabel = list(bgcolor = "white"))
```

### Overall Coverage 80

By Forecast Dates
```{r}
plot_cov80 <-
  plot_canonical(results, 
                 x = "forecast_date", 
                 y = "cov_80", 
                 aggr = mean,
                 grp_vars = c("forecaster","ahead"), 
                 facet_rows = "ahead") +
  labs(title = subtitle, x= "Forecast date", y = "Mean Coverage 80", color='Forecasters') +
  geom_point(aes(text = sprintf("Forecast Date: %s<br>Coverage: %s <br>Forecaster: %s", 
                                forecast_date, 
                                round(cov_80, digits = 2),
                                color)),
             alpha = 0.05) +
  facet_wrap(~ahead, nrow = 4, labeller = labeller(ahead = weeks.label)) +
  theme(strip.background = element_rect(fill = "white")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = "center")) +
  geom_line(mapping = aes(y = .8), )

if (params$colorblind_palette) {
  plot_cov80 <- plot_cov80 + 
    scale_color_viridis_d()
}

ggplotly(plot_cov80, tooltip="text", height=800, width=1000) %>% 
  layout(hoverlabel = list(bgcolor = "white"))
```

## Geometric mean relative WIS {.tabset}

### By Weeks Ahead

```{r by ahead,height=800, width= 1000, message = FALSE, warning = FALSE}
geom_mean <- function(x) prod(x)^(1/length(x))
#geom_mean <- exp(mean(log((x+1)/(y+1)))) #still need to figure this out

mean_wis <- 
  plot_canonical(results %>% 
                   filter(wis > 0), 
                 x = "ahead", 
                 y = "wis", 
                 aggr = geom_mean,
                 base_forecaster = "COVIDhub-trained_ensemble", 
                 scale_before_aggr = TRUE) + 
  labs(title = subtitle, 
       x = "Weeks Ahead", 
       y = "Geometric mean relative WIS", 
       color = "Forecasters") +
  geom_point(aes(text = sprintf("Weeks Ahead: %s<br>WIS: %s <br>Forecaster: %s", 
                                ahead, 
                                round(wis, digits = 2),
                                color)),
             alpha = 0.05) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = "center")) +
  geom_line(mapping = aes(y = 1))

if (params$colorblind_palette) {
  mean_wis <- mean_wis + 
    scale_color_viridis_d()
}

ggplotly(mean_wis, tooltip="text", width= 1000) %>% 
  layout(hoverlabel = list(bgcolor = "white"))
```

### By Forecast Date

```{r by date,height=800, width= 1000, message = FALSE, warning = FALSE}

mean_wis_forecast_date <- 
plot_canonical(results %>% 
                 filter(wis > 0), 
               x = "forecast_date", 
               y = "wis", 
               aggr = geom_mean, 
               facet_rows = "ahead",
               grp_vars = c("forecaster", "ahead"),
               base_forecaster = "COVIDhub-ensemble", 
               scale_before_aggr = TRUE) +
  theme(legend.position = "bottom") + 
  labs(title = subtitle, 
       x = "Forecast date", 
       y = "Geometric mean relative WIS", 
       color = "Forecasters") +
  geom_point(aes(text = sprintf("Forecast Date: %s<br>WIS: %s <br>Forecaster: %s", 
                                forecast_date, 
                                round(wis, digits = 2),
                                color)),
             alpha = 0.05) +
  facet_wrap(~ahead, nrow = 4, labeller = labeller(ahead = weeks.label)) +
  theme(strip.background = element_rect(fill = "white")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = "center")) +
  geom_line(mapping = aes(y = 1))

if (params$colorblind_palette) {
  mean_wis_forecast_date <- mean_wis_forecast_date + 
    scale_color_viridis_d()
}

ggplotly(mean_wis_forecast_date, tooltip = "text", height=800, width= 1000) %>% 
  layout(hoverlabel = list(bgcolor = "white"))
```

```{r, message = FALSE, warning = FALSE, include = FALSE}
## Scores by target date (not forecast date) 

# target_wis <-
#   plot_canonical(results, x = "target_end_date", y = "wis", aggr = mean,
#                dots = TRUE, grp_vars = "forecaster") + 
#   labs(title = subtitle, x = "Target date", y = "Mean WIS", color='Forecasters') +
#   scale_y_log10() +
#   geom_point(aes(text=sprintf("Target End Date: %s<br>WIS: %s <br>Forecaster: %s", target_end_date, round(wis, digits=2),color)),alpha = 0.05) +
#   theme_minimal() +
#   theme(plot.title = element_text(hjust = 'center'))
# 
# target_wis_ahead <-
#   plot_canonical(results, x = "target_end_date", y = "wis", aggr = mean,
#                dots = TRUE, grp_vars = c("forecaster", "ahead"), 
#                facet_rows = "ahead") +
#   labs(title = subtitle, x = "Target date", y = "Mean WIS", color='Forecasters') +
#   scale_y_log10() +
#   geom_point(aes(text=sprintf("Target End Date: %s<br>WIS: %s <br>Forecaster: %s", target_end_date, round(wis, digits=2),color)),alpha = 0.05)+
#   facet_wrap(~ahead, nrow = 4, labeller = labeller(ahead=weeks.label))+
#   theme(strip.background = element_rect(fill = "white")) +
#   theme_minimal() +
#   theme(plot.title = element_text(hjust = 'center'))
# 
# ggplotly(target_wis, tooltip="text", width= 1000) %>% layout(hoverlabel=list(bgcolor='white'))
# ggplotly(target_wis_ahead, tooltip="text", height=800, width= 1000) %>% layout(hoverlabel=list(bgcolor='white'))
```

## Maps {.tabset}

Note that the results are scaled by population.

```{r maps-processing, warning=FALSE}
library(sf)

results_intersect <- intersect_averagers(scores, c("forecaster"), c("forecast_date", "geo_value")) %>%
  select(c("ahead", "geo_value", "forecaster","forecast_date", "data_source", "signal","target_end_date","incidence_period","actual","wis","ae","cov_80"))

maps <- results_intersect %>%
  group_by(geo_value, forecaster) %>%
  summarise(across(wis:cov_80, mean)) %>%
  left_join(animalia::state_population, by = "geo_value") %>%
  mutate(across(wis:cov_80, ~ .x / population * 1e5)) %>%
  pivot_longer(wis:cov_80, names_to = "score") %>%
  group_by(score) %>%
  mutate(time_value = Sys.Date(),
         r = max(value)) %>%
  group_by(forecaster, .add = TRUE) %>%
  group_split()

# for county prediction, set geo_type = "county"
maps <- purrr::map(maps, 
                   ~as.covidcast_signal(
                     .x, signal = .x$score[1], 
                     data_source = .x$forecaster[1], 
                     geo_type = "state"))

maps <- purrr::map(maps,
                   ~plot(.x, 
                         choro_col = scales::viridis_pal()(3),
                         range = c(0,.x$r[1])))

nfcasts <- length(unique(results$forecaster))
```

### Mean AE

```{r map-ae, fig.width=12, fig.height=8}
# original code
cowplot::plot_grid(plotlist = maps[1:nfcasts], ncol = 3)
```

### Mean Coverage 80

```{r map-cov80, fig.width=12, fig.height=8}
cowplot::plot_grid(plotlist = maps[(nfcasts+1):(nfcasts*2)], ncol = 3)
```

### Mean WIS

```{r map-wis, fig.width=12, fig.height=8}
cowplot::plot_grid(plotlist = maps[((nfcasts*2)+1):length(maps)], ncol = 3)
```

<!-- ## Trajectory plots -->

<!-- For county predictions, you might want to change the `fig.height` and `fig.width` chunk options here (in other notebooks we use `fig.height = 120` and `fig.width = 30`). -->

<!-- ```{r get_predictions} -->
<!-- options(timeout=200) -->
<!-- url_pred <- "https://forecast-eval.s3.us-east-2.amazonaws.com/predictions_cards.rds" -->
<!-- download.file(url_pred, "predictions.RDS") # download to disk -->
<!-- state_predictions <- readRDS(paste0(here(), "/predictions.RDS")) -->

<!-- # Check how many aheads in the state_predictions data -->
<!-- # Check the forecast_end_date column -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # for county prediction, set geo_type = "county" -->
<!-- state.label = c(state.name, "Washington D.C.", "Porto Rico") -->
<!-- names(state.label) = c(tolower(state.abb), "dc", "pr") -->

<!-- # trajectory plots for selected forecaster -->
<!-- pd <- evalcast:::setup_plot_trajectory( -->
<!--   state_predictions %>% filter(forecaster=="COVIDhub-ensemble" & forecast_date %in% forecast_dates), -->
<!--   intervals = 0.8, -->
<!--   geo_type = "state", -->
<!--   start_day = min(forecast_dates) - 60) -->

<!-- g <- ggplot(pd$truth_df, mapping = aes(x = target_end_date)) -->
<!-- # build the fan -->
<!-- g <- g + geom_ribbon( -->
<!--   data = pd$quantiles_df %>% -->
<!--     mutate(upper = round(upper, 2), -->
<!--            lower = round(lower, 2)), -->
<!--   mapping = aes(ymin = lower,  -->
<!--                 ymax = upper,  -->
<!--                 fill = forecaster, -->
<!--                 group = interaction(forecaster, forecast_date)), -->
<!--   alpha = .1) + -->
<!--   scale_fill_viridis_d(begin=.15, end=.85) -->

<!-- # line layer -->
<!-- g <- g + -->
<!--   #geom_line(aes(y = .data$value.y), color = "#3182BD") + # corrected -->
<!--   geom_line(aes(y = value), size = .5) + # reported -->
<!--   geom_line(data = pd$points_df, -->
<!--             mapping = aes(y = value, -->
<!--                           color = forecaster,  -->
<!--                           group = interaction(forecaster, forecast_date)),  -->
<!--             size = .5) + -->
<!--   geom_point(aes(y = value), size = 1) + # reported gets dots -->
<!--   geom_point(data = pd$points_df %>% -->
<!--                mutate(value = round(value, 2)), -->
<!--              mapping = aes(y = value, color = forecaster), -->
<!--              size = 1) + -->
<!--   scale_color_viridis_d(begin=.15, end=.85) -->
<!-- ``` -->


<!-- ```{r state trajectories, fig.height = 60, fig.width = 20, dev="CairoSVG"} -->
<!-- states <- g  + -->
<!--   facet_wrap(~geo_value,  -->
<!--              scales = "free_y",  -->
<!--              ncol = 3,  -->
<!--              labeller = labeller(geo_value = state.label)) + -->
<!--   labs(x = "", y = "") + -->
<!--   theme_bw() +  -->
<!--   theme(legend.position = "none", strip.background = element_rect(fill = "white")) -->

<!-- ggplotly(states, height=5000, width= 1000) -->
<!-- ``` -->
